---
Title: "Analysis of the Adverse Health and Economic Impacts of US Storms"
Author: “Kris Tan”
date: “07 Aug 2018”
output: html_document
---

## 1: Synopsis
This analysis explores the U.S. National Oceanic and Atmospheric Administration's (NOAA) storm database describes the features of major weather events from 1950 to 2011.

The current report investigates on damages to:

1) Population health (injuries and fatalities);
2) Economical consequences of the affected communities. 

The analysis confirmed that excessive heat and tornado are most harmful to Population health, while flood, drought, and hurricane/typhoon have the greatest economical consequences.

Information on the Data: [Documentation](https://d396qusza40orc.cloudfront.net/repdata%2Fpeer2_doc%2Fpd01016005curr.pdf)

## Data Processing

### Dependencies

The following code is used to load the dependencies required for the data analysis.
```{r DataLoading}
#Read and initialize library and data
library('data.table')
library('dplyr')
library('ggplot2')
library('grid')
library('gridExtra')

StormData = read.csv("//Users/user/Documents/KrisTan/Coursera/StormData.csv")
summary(StormData)
```


# Types of Events that are most harmful to Population Health?

The analysis of fatalities and injuries according to different types of weather events.

Plot 1: Death Toll by Event Type

```{r, echo=TRUE}
StormData %>%
      select(FATALITIES, EVTYPE) %>%
      group_by(EVTYPE) %>%
      summarise(SumFATALITIES = sum(FATALITIES)) %>%
      top_n(n = 8, wt = SumFATALITIES) %>%
      ggplot(aes(y = SumFATALITIES, x = reorder(x = EVTYPE, X = SumFATALITIES), fill=EVTYPE))+
      geom_bar(stat = "identity", show.legend = FALSE) +
      #geom_text(aes(label=SumFATALITIES), size = 4, hjust = 0.5, vjust = -0.1) +
      xlab(label = "") +
      ylab(label = "Death Toll") +
      coord_flip() +
      theme_light()
```

Plot 2: Injuries by Event type

```{r, echo=TRUE}
StormData %>%
      select(INJURIES, EVTYPE) %>%
      group_by(EVTYPE) %>%
      summarise(SumINJURIES = sum(INJURIES)) %>%
      top_n(n = 8, wt = SumINJURIES) %>%
      ggplot(aes(y = SumINJURIES, x = reorder(x = EVTYPE, X = SumINJURIES), fill=EVTYPE))+
      geom_bar(stat = "identity", show.legend = FALSE) +
      #geom_text(aes(label=SumINJURIES), size = 4, hjust = 0.5, vjust = -0.1) +
      xlab(label = "") +
      ylab(label = "Injuries") +
      coord_flip() +
      theme_light()
```
# Types of events that have the greatest economical consequences?

Plot 3: Property Damage Estimates by Event Type

```{r, echo=TRUE}
StormData %>%
      select(PROPDMG, PROPDMGEXP2, EVTYPE) %>%
      group_by(EVTYPE) %>%
      mutate(SumPROPDMGEXP = (PROPDMG * PROPDMGEXP2)) %>%
      summarise(SumPROPDMGEXP2 = sum(SumPROPDMGEXP)) %>%
      top_n(n = 8, wt = SumPROPDMGEXP2) %>%
      ggplot(aes(y = SumPROPDMGEXP2, x = reorder(x = EVTYPE, X = SumPROPDMGEXP2), fill=EVTYPE))+
      geom_bar(stat = "identity", show.legend = FALSE) +
      #geom_text(aes(label=SumFATALITIES), size = 4, hjust = 0.5, vjust = -0.1) +
      xlab(label = "") +
      ylab(label = "Property Damage Estimates") +
      coord_flip() +
      theme_light()

```
# Results

#### According to the analysis:
1) Tornado is the most harmful to population health.
2) Flood has the greatest economical consequences. 

